### 2025-04-13 project3/draft.tex
11

In the evolving landscape of financial markets, insurance products such as VAs have gained a substantial amount of interest due to their ability to provide both investment growth and guaranteed benefits. 
Managing risks associated with these products in volatile market conditions is a complex task that demands sophisticated financial modeling techniques such as nested simulation.
A standard nested simulation procedure involves running a large number of simulations to generate a dataset of scenario-wise contract losses with two levels of simulations.
Practitioners can estimate tail risk measures using the realized losses for different risk scenarios obtained from these simulations.
Nested simulation is extremely computationally expensive.
To address the computational burden, metamodeling techniques have been proposed, where a surrogate model approximates the outcomes of a complex simulation procedure. 
Traditional machine learning models often struggle to capture the intricate nonlinear relationships and temporal dependencies inherent in financial data.
In particular, DNNs, and specifically LSTM networks, have been employed as metamodels to predict the outcomes of the inner simulations efficiently.
Chapter~\ref{chap:project2} introduced a metamodeling-based nested simulation framework for dynamic hedging of VAs that approximates the inner simulation by an LSTM metamodel.
It shows that LSTMs, as a type of gated RNNs, were well-suited as metamodels for Monte Carlo simulation of financial time series and could capture the long-term dependencies in the simulation dataset.
This property is crucial for the application of dynamic hedging of VAs.

Despite the advantages of using DNN metamodels, computational challenges arise when market conditions change or new VA contracts with different features are introduced. 
Retraining neural network metamodels from the scratch in response to every change is computationally inefficient and time-consuming.
Practitioners with limited resources face several challenges:
\begin{enumerate}
    \item The simulation budget is limited, but DNN metamodels require extensive training data to learn the feature-label relationship~\citep{golestaneh2024many}. 
    \item The computational cost of retraining a suitable DNN metamodel is often much higher than fitting a traditional parametric regression.
    \item Financial markets are dynamic with frequent shifts in volatility, interest rates, and other risk factors~\citep{cont2001empirical}.
\end{enumerate}

Constraints on the simulation budget is the most critical challenge for dynamic hedging of VAs.
Any method that involves nested simulation is computationally expensive, and the computational cost is often prohibitive for changing market conditions.
Therefore, it is essential to develop methods that can rapidly adapt existing metamodels to new conditions without incurring the full computational cost of retraining and extensive addtional simulations.
Transfer Learning (TL) offers a compelling solution to this problem by enabling the reuse of a pre-trained model on a new but related task~\citep{pan2009survey}.
In the context of DNNs, TL involves leveraging the knowledge acquired during training on one dataset to improve learning performance on a different dataset~\citep{yosinski2014transferable}.
This approach can substantially reduce the amount of training data needed, training time and computational resources while enhancing model generalization.
Instead of starting from the scratch, a new DNN metamodel can be built on top of a pre-trained metamodel and fine-tuned on new data.
This allows the metamodel to adapt quickly to unseen conditions and new VA contracts.

In this chapter, we explore the application of TL to the dynamic hedging of VAs using LSTM metamodels.
We propose a novel TL framework that accelerates the training of DNN metamodels for nested simulation procedures in dynamic hedging.
This setting is particularly relevant for dynamic hedging problems, as insurance companies often underwrite new contracts under changing market conditions and actuarial assumptions.
This chapter proposes a TL framework that makes fast adaptation of metamodels possible.
Our approach involves pre-training an LSTM network on a large dataset of VA simulations and then fine-tuning it on a smaller dataset of new VA contracts.
We evaluate the performance of the TL framework on a real-world dataset of VA contracts and compare it with training a LSTM metamodel from scratch.
More importantly, we show the strengths and weaknesses of different TL techniques for rapid adaptation of LSTM metamodels.

The rest of this chapter is organized as follows.
Section~\ref{sec3:background} provides an overview of the dynamic hedging problem for VAs and the use of LSTM networks as metamodels in a nested simulation procedure.
Section~\ref{sec3:transfer_learning} introduces the TL framework for rapid adaptation of LSTM metamodels in dynamic hedging.
Section~\ref{sec3:experiments} presents the experimental setup and results, comparing the performance of TL with training from scratch.
Finally, Section~\ref{sec3:conclusion} concludes the chapter and discusses future research directions.

@@

TL is a machine learning paradigm where knowledge acquired from a source task is utilized to improve learning performance on a related target task.
One of the primary applications of TL in finance is asset price prediction. 
Traditional models, such as geometric Brownian motion (GBM), autoregressive integrated moving average (ARIMA) and generalized autoregressive conditional heteroskedasticity (GARCH), have been widely used in time series modeling.
ARIMA models~\citep{box1970distribution} combine autoregression, differencing, and moving average components to capture linear relationships in time series data.
GARCH models~\citep{bollerslev1990modelling}, extend ARCH models to better capture volatility clustering, which is a key stylized fact of financial time series~\citep{cont2001empirical}.
However, these models often struggle to capture complex dynamics, e.g., regime changes~\citep{hamilton1989new} and extreme market events~\citep{embrechts2013modelling} that are prevalent in financial markets.
In Chapter~\ref{chap:project2}, our numerical experiment uses a regime-switching geometric Brownian motion (RS-GBM) as the underlying asset model.
For estimating tail risk measures, a metamodel-based nested simulation procedure is proposed to approximate the inner simulation.
We have shown DNN metamodels like RNN and LSTM networks have demonstrated substantial improvements in modeling temporal dependencies, and they save a substantial amount of simulation budget.
However, training these models from the scratch requires an extensive amount of simulation data, which may not always be available for specific assets or under certain market conditions.

TL offers a solution to this problem by leveraging knowledge from related assets or tasks to improve the learning performance of DNN metamodels on the target task.
In algorithmic trading,~\cite{jeong2019improving} used TL to enhance the performance of a reinforcement learning agent by preventing overfitting from an insufficient amount of market data.
TL techniques have also been applied in building fraud detection systems. 
Financial fraud often exhibits subtle and evolving patterns, and it is challenging to develop robust detection models.
By transferring previous knowledge from detected fraud cases, models can adapt to detect new fraud schemes more effectively~\citep{lebichot2021transfer}.
\cite{yan2024comprehensive} conduct a comprehensive survey study of current TL techniques in financial applications, and they find almost all applications of TL only employ parameter transfer, where the pre-trained model is fine-tuned on the target task.
Our multi-task learning framework in Section~\ref{sec3:transfer_learning} extends beyond parameter transfer to shared representation learning across multiple tasks.

Regarding a nested simulation of VAs,~\cite{cheng2019fast} is the most relevant study to our work, where they proposed a TL framework for fast valuation of large portfolios of VAs.
Instead of using stochastic kriging~\citep{gan2015valuation}, they employed a pre-trained DNN to select the best representative scenarios from a large portfolio of VAs.
In our work, we focus on the application of TL to accelerate the training of LSTM metamodels for a nested simulation in dynamic hedging of VAs.
The fine-tuned LSTM metamodels can be readily adapted to the two-stage procedure and the single-stage procedure in Chapter~\ref{chap:project2} to predict the contract losses under different scenarios.

In our context of DNNs metamodeling-based simulation procedures for hedging VAs, TL involves pre-training a DNN metamodel where the simulation budget is abundant and then fine-tuning it on a smaller dataset of new contracts or market conditions.
Written on the same underlying asset, different VA contracts may share common features or exhibit similar patterns, especially temporal dependencies and regime changes in the underlying financial time series.
Similarly, two VAs with the same features but based on different underlying assets may have some shared characteristics that can be leveraged to improve the learning performance of the metamodel.
By transferring knowledge from a pre-trained model to a new but related task, TL can substantially reduce the computational cost of training the metamodel on the target task.

LSTM networks are well-suited for modeling sequential data due to their ability to capture long-term dependencies~\citep{hochreiter1997long}.
For the application of VA risk management using metamodel-based nested simulation, LSTM networks approximate the inner simulation, i.e., the mapping from scenarios to the scenario-wise contract losses.
In Chapter~\ref{chap:project2}, we treat metamodeling as a supervised learning problem and demonstrated that LSTM networks could effectively model this complex relationships with extensive training on a large dataset of VA simulations.
The total computational cost originates from two sources: running the standard nested simulation procedure to generate the training data and training the LSTM network on the generate dataset.
Given a new VA contract, these steps need to be repeated to adapt the LSTM metamodel to new market conditions and contract features.
For a DNN with a large number of parameters, retraining the LSTM network from the scratch can be computationally expensive.
Generating a large training dataset using the standard nested simulation procedure is costly, and the computational burden can be prohibitive for certain assets or market conditions.
TL offers an alternative approach that keeps the previous knowledge, i.e., the pre-trained model, as a foundation.
Building on the pre-trained model, the LSTM network can be fine-tuned on a smaller dataset of new contracts or market conditions.
This approach accelerates the adaptation of the LSTM metamodel to new conditions.
Computational savings come in two forms: 
\begin{enumerate}
    \item   fine-tuning requires less training time than training LSTM metamodels from scratch, and
    \item   fewer training data points are needed to achieve a good LSTM metamodel, which is particularly beneficial when the standard nested simulation procedure is costly to implement.
\end{enumerate}

@@

A typical TL framework for supervised learning consists of the following:
\begin{itemize}
    \item   \textbf{source domain}: $\mathcal{D}_{\text{So}} = \{\mathcal{X}_{\text{So}}, F_{\text{So}}(\mathbf{X})\}$,
    \item   \textbf{source task}: $\mathcal{T}_{\text{So}} = \{\mathcal{Y}_{\text{So}}, f_{\text{So}}(\cdot)\}$,
    \item   \textbf{target domain}: $\mathcal{D}_{\text{Ta}} = \{\mathcal{X}_{\text{Ta}}, F_{\text{Ta}}(\cdot)\}$,
    \item   \textbf{target task}: $\mathcal{T}_{\text{Ta}} = \{\mathcal{Y}_{\text{Ta}}, f_{\text{Ta}}(\cdot)\}$,
\end{itemize}
where $\mathcal{X}_{\text{So}}$ and $\mathcal{X}_{\text{Ta}}$ include input features derived from the outer simulation of two different VA contracts or market conditions, and $F_{\text{So}}$ and $F_{\text{Ta}}$ are the marginal probability distributions of the source and target domains, respectively.
Hence, the source domain and task are related to one VA contract for a certain market condition, while the target domain and task are related to new VA contracts and market conditions for which we don't have the simulation budget to run the standard nested simulation procedure extensively.
In our numerical experiments, the dataset is generated by running the standard nested simulation procedure in Algorithm~\ref{alg2:standardProcedure} for a large number of VA contracts.
The input features $\mathbf{X}$ are the return vectors from Equation~\eqref{eq2:return}, and the output labels $L$ are the contract losses for different outer scenarios.
The stock price vector $\mathbf{S}$ is not crucial for the discussion of this chapter and is omitted since it can be immediately recovered from $\mathbf{X}$ and the initial stock price $S_0$.
We keep the number of features to be $\num{240}$ for both source and target domains as in Section~\ref{sec2:numerical}.
The predictive function $f_{\text{So}}$ and $f_{\text{Ta}}$ are trained to estimate outputs $L_{\text{So}}$ and $L_{\text{Ta}}$, i.e., approximating the inner simulations under the source and target tasks, respectively.
TL seeks to improve the learning of the target predictive function $f_{\text{Ta}}(\cdot)$ in $\mathcal{D}_{\text{Ta}}$ by leveraging the previous training on $\mathcal{D}_{\text{So}}$ and $f_{\text{So}}(\cdot)$, particularly when $\mathcal{D}_{\text{So}} \neq \mathcal{D}_{\text{Ta}}$ or $\mathcal{T}_{\text{So}} \neq \mathcal{T}_{\text{Ta}}$.
In Chapter~\ref{chap:project2}, we have $\mathcal{D}_{\text{So}} = \mathcal{D}_{\text{Ta}}$ and $\mathcal{T}_{\text{So}} = \mathcal{T}_{\text{Ta}}$.

In our TL setting, both the source and target tasks involve learning a mapping from the risk factors to the VA contract losses.
For the source task, we have a dataset $\mathcal{D}_{\text{So}} = { (X_{\text{So}}^{(i)}, L_{\text{So}}^{(i)}) }_{i=1}^{M_{\text{So}}}$, where $M_{\text{So}}$ is the number of training samples, i.e., the number of outer simulation paths used to generate the dataset with a standard nested simulation procedure.
$X_{\text{So}}^{(i)} \in \mathcal{X}_{\text{So}}$ and $L_{\text{So}}^{(i)} \in \mathcal{Y}_{\text{So}}$.
The ultimate objective is to learn a metamodel $f_{\text{Ta}}$ that predicts the VA contract losses $L_{\text{So}}$ from the scenarios $X_{\text{Ta}}$ in the form of a financial time series.
TL starts by training a metamodel $f_{\text{So}}$ on the source domain $\mathcal{D}_{\text{So}}$.
For both the two-stage procedure and the single-stage procedure in Chapter~\ref{chap:project2}, $f_{\text{So}}$ is approximated by an LSTM network $f_{\text{So}}(\cdot ; \theta_{\text{So}})$, where the parameters of the LSTM network, $\theta_{\text{So}}$ are learned by minimizing an MSE loss function on the source domain.
Then the pre-trained parameters $\theta_{\text{So}}$ to inform the learning of $f_{\text{Ta}}(\cdot ; \theta_{\text{Ta}})$ on the target domain $\mathcal{D}_{\text{Ta}}$.
The training on the target domain should encourage similarity between $\theta_{\text{So}}$ and $\theta_{\text{Ta}}$ to facilitate the transfer of knowledge.

The most common TL techniques for supervised learning include fine-tuning, layer freezing, and multi-task learning. 
These techniques can be categorized based on how they leverage pre-trained metamodels and how they encourage the similarity between $f_{\text{So}}(\cdot ; \theta_{\text{So}})$ and $f_{\text{Ta}}(\cdot ; \theta_{\text{Ta}})$.

@@

Fine-tuning is a commonly used TL technique that uses the \textbf{same neural network architecture} for both the source and target tasks.
We first train the source LSTM metamodel $f_{\text{So}}(\cdot; \theta_{\text{So}})$ on $\mathcal{D}_{\text{So}}$, capturing temporal dependencies and patterns relevant to the source task. 
For the target task, we initialize $\theta_{\text{Ta}} = \theta_{\text{So}}$ and proceed to fine-tune the entire network on $\mathcal{D}_{\text{Ta}}$ by using a smaller learning rate.

@@

    \STATE Train a LSTM metamodel $f_{\text{So}}(\cdot; \theta_{\text{So}})$ on $\mathcal{D}_{\text{So}}$:
    \begin{equation}
        \theta_{\text{So}} = \min_{\theta} \frac{1}{M_{\text{So}}} \sum_{i=1}^{M_{\text{So}}} \left( f_{\text{So}}(X_{\text{So}}^{(i)}; \theta) - L_{\text{So}}^{(i)} \right)^2.
    \end{equation}

@@

    \STATE Fine-tune the entire LSTM model $f_{\text{Ta}}(\cdot; \theta_{\text{Ta}})$ on the target dataset $\mathcal{D}_{\text{Ta}}$ using a smaller learning rate $\alpha_{\text{Ta}}$:
    \begin{equation}
        \theta_{\text{Ta}} = \min_{\theta} \frac{1}{M_{\text{Ta}}} \sum_{i=1}^{M_{\text{Ta}}} \left( f_{\text{Ta}}(X_{\text{Ta}}^{(i)}; \theta) - L_{\text{Ta}}^{(i)} \right)^2.
    \end{equation}

    \STATE \textbf{Output:} Final adapted LSTM metamodel $f_{\text{Ta}}(\cdot; \theta_{\text{Ta}})$ for the target task.
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg3:fineTuning} assumes that learned sequential representations are beneficial for the target task.
Only minor adjustments are needed to adapt the metamodel to the new task, and the training process is accelerated by the pre-trained parameters.
Fine-tuning is particularly useful when the nested simulation procedures for the two VA contracts are closely related.
Only minor adjustments are needed to adapt the LSTM metamodel to the new domain.

@@

        \STATE Train LSTM model $f_{\text{So}}(\cdot; \theta_{\text{So}})$ on $\mathcal{D}_{\text{So}}$:
        \begin{equation}
            \theta_{\text{So}} = [\theta_0, \theta_1] = \min_{\theta} \frac{1}{M_{\text{So}}} \sum_{i=1}^{M_{\text{So}}} \left( f_{\text{So}}(X_{\text{So}}^{(i)}; \theta) - L_{\text{So}}^{(i)} \right)^2.
        \end{equation}

@@

        \STATE \textbf{Output:} Final adapted LSTM metamodel $f_{\text{Ta}}(\cdot; [\theta_0, \theta_1])$ for the target task.
    \end{algorithmic}
    \end{algorithm}

@@

For the implementation of our metamodeling studied in Chapter~\ref{chap:project2}, we consider a multi-task learning framework where the LSTM layers are shared across multiple tasks, and each task has its own fully connected layer for prediction.
The network parameters are divided into shared parameters $\theta_0$ (LSTM layers) and task-specific parameters $\theta_k$ (fully connected layers for task $k$). 
As opposed to layer freezing, all parameters are trainable.

@@

        \STATE \textbf{Output:} Trained multi-task LSTM metamodel $f(\cdot; \theta_0, \{\theta_k\}_{k=1}^K)$ for all $K$ tasks.
    \end{algorithmic}
\end{algorithm}

>>>>

In the evolving landscape of financial markets, insurance products such as \gls{va}s have gained a substantial amount of interest due to their ability to provide both investment growth and guaranteed benefits. 
Managing risks associated with these products in volatile market conditions is a complex task that demands sophisticated financial modeling techniques such as nested simulation.
A standard nested simulation procedure involves running a large number of simulations to generate a dataset of scenario-wise contract losses with two levels of simulations.
Practitioners can estimate tail risk measures using the realized losses for different risk scenarios obtained from these simulations.
Nested simulation is extremely computationally expensive.
To address the computational burden, metamodeling techniques have been proposed, where a surrogate model approximates the outcomes of a complex simulation procedure. 
Traditional \gls{ml} models often struggle to capture the intricate nonlinear relationships and temporal dependencies inherent in financial data.
In particular, \gls{dnn}s, and specifically \gls{lstm} networks, have been employed as metamodels to predict the outcomes of the inner simulations efficiently.
Chapter~\ref{chap:project2} introduced a metamodeling-based nested simulation framework for dynamic hedging of \gls{va}s that approximates the inner simulation by an \gls{lstm} metamodel.
It shows that \gls{lstm}s, as a type of gated \gls{rnn}s, were well-suited as metamodels for \gls{mc} simulation of financial time series and could capture the long-term dependencies in the simulation dataset.
This property is crucial for the application of dynamic hedging of \gls{va}s.

Despite the advantages of using \gls{dnn} metamodels, computational challenges arise when market conditions change or new \gls{va} contracts with different features are introduced. 
Retraining neural network metamodels from the scratch in response to every change is computationally inefficient and time-consuming.
Practitioners with limited resources face several challenges:
\begin{enumerate}
    \item The simulation budget is limited, but \gls{dnn} metamodels require extensive training data to learn the feature-label relationship~\citep{golestaneh2024many}. 
    \item The computational cost of retraining a suitable \gls{dnn} metamodel is often much higher than fitting a traditional parametric regression.
    \item Financial markets are dynamic with frequent shifts in volatility, interest rates, and other risk factors~\citep{cont2001empirical}.
\end{enumerate}

Constraints on the simulation budget is the most critical challenge for dynamic hedging of \gls{va}s.
Any method that involves nested simulation is computationally expensive, and the computational cost is often prohibitive for changing market conditions.
Therefore, it is essential to develop methods that can rapidly adapt existing metamodels to new conditions without incurring the full computational cost of retraining and extensive addtional simulations.
Transfer Learning (\gls{tl}) offers a compelling solution to this problem by enabling the reuse of a pre-trained model on a new but related task~\citep{pan2009survey}.
In the context of \gls{dnn}s, \gls{tl} involves leveraging the knowledge acquired during training on one dataset to improve learning performance on a different dataset~\citep{yosinski2014transferable}.
This approach can substantially reduce the amount of training data needed, training time and computational resources while enhancing model generalization.
Instead of starting from the scratch, a new \gls{dnn} metamodel can be built on top of a pre-trained metamodel and fine-tuned on new data.
This allows the metamodel to adapt quickly to unseen conditions and new \gls{va} contracts.

In this chapter, we explore the application of \gls{tl} to the dynamic hedging of \gls{va}s using \gls{lstm} metamodels.
We propose a novel \gls{tl} framework that accelerates the training of \gls{dnn} metamodels for nested simulation procedures in dynamic hedging.
This setting is particularly relevant for dynamic hedging problems, as insurance companies often underwrite new contracts under changing market conditions and actuarial assumptions.
This chapter proposes a \gls{tl} framework that makes fast adaptation of metamodels possible.
Our approach involves pre-training an \gls{lstm} network on a large dataset of \gls{va} simulations and then fine-tuning it on a smaller dataset of new \gls{va} contracts.
We evaluate the performance of the \gls{tl} framework on a real-world dataset of \gls{va} contracts and compare it with training a \gls{lstm} metamodel from scratch.
More importantly, we show the strengths and weaknesses of different \gls{tl} techniques for rapid adaptation of \gls{lstm} metamodels.

The rest of this chapter is organized as follows.
Section~\ref{sec3:background} provides an overview of the dynamic hedging problem for \gls{va}s and the use of \gls{lstm} networks as metamodels in a nested simulation procedure.
Section~\ref{sec3:transfer_learning} introduces the \gls{tl} framework for rapid adaptation of \gls{lstm} metamodels in dynamic hedging.
Section~\ref{sec3:experiments} presents the experimental setup and results, comparing the performance of \gls{tl} with training from scratch.
Finally, Section~\ref{sec3:conclusion} concludes the chapter and discusses future research directions.

@@

\gls{tl} is a machine learning paradigm where knowledge acquired from a source task is utilized to improve learning performance on a related target task.
One of the primary applications of \gls{tl} in finance is asset price prediction. 
Traditional models, such as geometric Brownian motion (\gls{gbm}), autoregressive integrated moving average (\gls{arima}) and generalized autoregressive conditional heteroskedasticity (\gls{garch}), have been widely used in time series modeling.
\gls{arima} models~\citep{box1970distribution} combine autoregression, differencing, and moving average components to capture linear relationships in time series data.
\gls{garch} models~\citep{bollerslev1990modelling}, extend \gls{arch} models to better capture volatility clustering, which is a key stylized fact of financial time series~\citep{cont2001empirical}.
However, these models often struggle to capture complex dynamics, e.g., regime changes~\citep{hamilton1989new} and extreme market events~\citep{embrechts2013modelling} that are prevalent in financial markets.
In Chapter~\ref{chap:project2}, our numerical experiment uses a regime-switching geometric Brownian motion (\gls{rsgbm}) as the underlying asset model.
For estimating tail risk measures, a metamodel-based nested simulation procedure is proposed to approximate the inner simulation.
We have shown \gls{dnn} metamodels like \gls{rnn} and \gls{lstm} networks have demonstrated substantial improvements in modeling temporal dependencies, and they save a substantial amount of simulation budget.
However, training these models from the scratch requires an extensive amount of simulation data, which may not always be available for specific assets or under certain market conditions.

\gls{tl} offers a solution to this problem by leveraging knowledge from related assets or tasks to improve the learning performance of \gls{dnn} metamodels on the target task.
In algorithmic trading,~\cite{jeong2019improving} used \gls{tl} to enhance the performance of a reinforcement learning agent by preventing overfitting from an insufficient amount of market data.
\gls{tl} techniques have also been applied in building fraud detection systems. 
Financial fraud often exhibits subtle and evolving patterns, and it is challenging to develop robust detection models.
By transferring previous knowledge from detected fraud cases, models can adapt to detect new fraud schemes more effectively~\citep{lebichot2021transfer}.
\cite{yan2024comprehensive} conduct a comprehensive survey study of current \gls{tl} techniques in financial applications, and they find almost all applications of \gls{tl} only employ parameter transfer, where the pre-trained model is fine-tuned on the target task.
Our multi-task learning framework in Section~\ref{sec3:transfer_learning} extends beyond parameter transfer to shared representation learning across multiple tasks.

Regarding a nested simulation of \gls{va}s,~\cite{cheng2019fast} is the most relevant study to our work, where they proposed a \gls{tl} framework for fast valuation of large portfolios of \gls{va}s.
Instead of using stochastic kriging~\citep{gan2015valuation}, they employed a pre-trained \gls{dnn} to select the best representative scenarios from a large portfolio of \gls{va}s.
In our work, we focus on the application of \gls{tl} to accelerate the training of \gls{lstm} metamodels for a nested simulation in dynamic hedging of \gls{va}s.
The fine-tuned \gls{lstm} metamodels can be readily adapted to the two-stage procedure and the single-stage procedure in Chapter~\ref{chap:project2} to predict the contract losses under different scenarios.

In our context of \gls{dnn}s metamodeling-based simulation procedures for hedging \gls{va}s, \gls{tl} involves pre-training a \gls{dnn} metamodel where the simulation budget is abundant and then fine-tuning it on a smaller dataset of new contracts or market conditions.
Written on the same underlying asset, different \gls{va} contracts may share common features or exhibit similar patterns, especially temporal dependencies and regime changes in the underlying financial time series.
Similarly, two \gls{va}s with the same features but based on different underlying assets may have some shared characteristics that can be leveraged to improve the learning performance of the metamodel.
By transferring knowledge from a pre-trained model to a new but related task, \gls{tl} can substantially reduce the computational cost of training the metamodel on the target task.

\gls{lstm} networks are well-suited for modeling sequential data due to their ability to capture long-term dependencies~\citep{hochreiter1997long}.
For the application of \gls{va} risk management using metamodel-based nested simulation, \gls{lstm} networks approximate the inner simulation, i.e., the mapping from scenarios to the scenario-wise contract losses.
In Chapter~\ref{chap:project2}, we treat metamodeling as a supervised learning problem and demonstrated that \gls{lstm} networks could effectively model this complex relationships with extensive training on a large dataset of \gls{va} simulations.
The total computational cost originates from two sources: running the standard nested simulation procedure to generate the training data and training the \gls{lstm} network on the generate dataset.
Given a new \gls{va} contract, these steps need to be repeated to adapt the \gls{lstm} metamodel to new market conditions and contract features.
For a \gls{dnn} with a large number of parameters, retraining the \gls{lstm} network from the scratch can be computationally expensive.
Generating a large training dataset using the standard nested simulation procedure is costly, and the computational burden can be prohibitive for certain assets or market conditions.
\gls{tl} offers an alternative approach that keeps the previous knowledge, i.e., the pre-trained model, as a foundation.
Building on the pre-trained model, the \gls{lstm} network can be fine-tuned on a smaller dataset of new contracts or market conditions.
This approach accelerates the adaptation of the \gls{lstm} metamodel to new conditions.
Computational savings come in two forms: 
\begin{enumerate}
    \item   fine-tuning requires less training time than training \gls{lstm} metamodels from scratch, and
    \item   fewer training data points are needed to achieve a good \gls{lstm} metamodel, which is particularly beneficial when the standard nested simulation procedure is costly to implement.
\end{enumerate}

@@

A typical \gls{tl} framework for supervised learning consists of the following:
\begin{itemize}
    \item   \textbf{source domain}: $\mathcal{D}_{\text{So}} = \{\mathcal{X}_{\text{So}}, F_{\text{So}}(\mathbf{X})\}$,
    \item   \textbf{source task}: $\mathcal{T}_{\text{So}} = \{\mathcal{Y}_{\text{So}}, f_{\text{So}}(\cdot)\}$,
    \item   \textbf{target domain}: $\mathcal{D}_{\text{Ta}} = \{\mathcal{X}_{\text{Ta}}, F_{\text{Ta}}(\cdot)\}$,
    \item   \textbf{target task}: $\mathcal{T}_{\text{Ta}} = \{\mathcal{Y}_{\text{Ta}}, f_{\text{Ta}}(\cdot)\}$,
\end{itemize}
where $\mathcal{X}_{\text{So}}$ and $\mathcal{X}_{\text{Ta}}$ include input features derived from the outer simulation of two different \gls{va} contracts or market conditions, and $F_{\text{So}}$ and $F_{\text{Ta}}$ are the marginal probability distributions of the source and target domains, respectively.
Hence, the source domain and task are related to one \gls{va} contract for a certain market condition, while the target domain and task are related to new \gls{va} contracts and market conditions for which we don't have the simulation budget to run the standard nested simulation procedure extensively.
In our numerical experiments, the dataset is generated by running the standard nested simulation procedure in Algorithm~\ref{alg2:standardProcedure} for a large number of \gls{va} contracts.
The input features $\mathbf{X}$ are the return vectors from Equation~\eqref{eq2:return}, and the output labels $L$ are the contract losses for different outer scenarios.
The stock price vector $\mathbf{S}$ is not crucial for the discussion of this chapter and is omitted since it can be immediately recovered from $\mathbf{X}$ and the initial stock price $S_0$.
We keep the number of features to be $\num{240}$ for both source and target domains as in Section~\ref{sec2:numerical}.
The predictive function $f_{\text{So}}$ and $f_{\text{Ta}}$ are trained to estimate outputs $L_{\text{So}}$ and $L_{\text{Ta}}$, i.e., approximating the inner simulations under the source and target tasks, respectively.
\gls{tl} seeks to improve the learning of the target predictive function $f_{\text{Ta}}(\cdot)$ in $\mathcal{D}_{\text{Ta}}$ by leveraging the previous training on $\mathcal{D}_{\text{So}}$ and $f_{\text{So}}(\cdot)$, particularly when $\mathcal{D}_{\text{So}} \neq \mathcal{D}_{\text{Ta}}$ or $\mathcal{T}_{\text{So}} \neq \mathcal{T}_{\text{Ta}}$.
In Chapter~\ref{chap:project2}, we have $\mathcal{D}_{\text{So}} = \mathcal{D}_{\text{Ta}}$ and $\mathcal{T}_{\text{So}} = \mathcal{T}_{\text{Ta}}$.

In our \gls{tl} setting, both the source and target tasks involve learning a mapping from the risk factors to the \gls{va} contract losses.
For the source task, we have a dataset $\mathcal{D}_{\text{So}} = { (X_{\text{So}}^{(i)}, L_{\text{So}}^{(i)}) }_{i=1}^{M_{\text{So}}}$, where $M_{\text{So}}$ is the number of training samples, i.e., the number of outer simulation paths used to generate the dataset with a standard nested simulation procedure.
$X_{\text{So}}^{(i)} \in \mathcal{X}_{\text{So}}$ and $L_{\text{So}}^{(i)} \in \mathcal{Y}_{\text{So}}$.
The ultimate objective is to learn a metamodel $f_{\text{Ta}}$ that predicts the \gls{va} contract losses $L_{\text{So}}$ from the scenarios $X_{\text{Ta}}$ in the form of a financial time series.
\gls{tl} starts by training a metamodel $f_{\text{So}}$ on the source domain $\mathcal{D}_{\text{So}}$.
For both the two-stage procedure and the single-stage procedure in Chapter~\ref{chap:project2}, $f_{\text{So}}$ is approximated by an \gls{lstm} network $f_{\text{So}}(\cdot ; \theta_{\text{So}})$, where the parameters of the \gls{lstm} network, $\theta_{\text{So}}$ are learned by minimizing an \gls{mse} loss function on the source domain.
Then the pre-trained parameters $\theta_{\text{So}}$ to inform the learning of $f_{\text{Ta}}(\cdot ; \theta_{\text{Ta}})$ on the target domain $\mathcal{D}_{\text{Ta}}$.
The training on the target domain should encourage similarity between $\theta_{\text{So}}$ and $\theta_{\text{Ta}}$ to facilitate the transfer of knowledge.

The most common \gls{tl} techniques for supervised learning include fine-tuning, layer freezing, and multi-task learning. 
These techniques can be categorized based on how they leverage pre-trained metamodels and how they encourage the similarity between $f_{\text{So}}(\cdot ; \theta_{\text{So}})$ and $f_{\text{Ta}}(\cdot ; \theta_{\text{Ta}})$.

@@

Fine-tuning is a commonly used \gls{tl} technique that uses the \textbf{same neural network architecture} for both the source and target tasks.
We first train the source \gls{lstm} metamodel $f_{\text{So}}(\cdot; \theta_{\text{So}})$ on $\mathcal{D}_{\text{So}}$, capturing temporal dependencies and patterns relevant to the source task. 
For the target task, we initialize $\theta_{\text{Ta}} = \theta_{\text{So}}$ and proceed to fine-tune the entire network on $\mathcal{D}_{\text{Ta}}$ by using a smaller learning rate.

@@

    \STATE Train a \gls{lstm} metamodel $f_{\text{So}}(\cdot; \theta_{\text{So}})$ on $\mathcal{D}_{\text{So}}$:
    \begin{equation}
        \theta_{\text{So}} = \min_{\theta} \frac{1}{M_{\text{So}}} \sum_{i=1}^{M_{\text{So}}} \left( f_{\text{So}}(X_{\text{So}}^{(i)}; \theta) - L_{\text{So}}^{(i)} \right)^2.
    \end{equation}

@@

    \STATE Fine-tune the entire \gls{lstm} model $f_{\text{Ta}}(\cdot; \theta_{\text{Ta}})$ on the target dataset $\mathcal{D}_{\text{Ta}}$ using a smaller learning rate $\alpha_{\text{Ta}}$:
    \begin{equation}
        \theta_{\text{Ta}} = \min_{\theta} \frac{1}{M_{\text{Ta}}} \sum_{i=1}^{M_{\text{Ta}}} \left( f_{\text{Ta}}(X_{\text{Ta}}^{(i)}; \theta) - L_{\text{Ta}}^{(i)} \right)^2.
    \end{equation}

    \STATE \textbf{Output:} Final adapted \gls{lstm} metamodel $f_{\text{Ta}}(\cdot; \theta_{\text{Ta}})$ for the target task.
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg3:fineTuning} assumes that learned sequential representations are beneficial for the target task.
Only minor adjustments are needed to adapt the metamodel to the new task, and the training process is accelerated by the pre-trained parameters.
Fine-tuning is particularly useful when the nested simulation procedures for the two \gls{va} contracts are closely related.
Only minor adjustments are needed to adapt the \gls{lstm} metamodel to the new domain.

@@

        \STATE Train \gls{lstm} model $f_{\text{So}}(\cdot; \theta_{\text{So}})$ on $\mathcal{D}_{\text{So}}$:
        \begin{equation}
            \theta_{\text{So}} = [\theta_0, \theta_1] = \min_{\theta} \frac{1}{M_{\text{So}}} \sum_{i=1}^{M_{\text{So}}} \left( f_{\text{So}}(X_{\text{So}}^{(i)}; \theta) - L_{\text{So}}^{(i)} \right)^2.
        \end{equation}

@@

        \STATE \textbf{Output:} Final adapted \gls{lstm} metamodel $f_{\text{Ta}}(\cdot; [\theta_0, \theta_1])$ for the target task.
    \end{algorithmic}
    \end{algorithm}

@@

For the implementation of our metamodeling studied in Chapter~\ref{chap:project2}, we consider a multi-task learning framework where the \gls{lstm} layers are shared across multiple tasks, and each task has its own fully connected layer for prediction.
The network parameters are divided into shared parameters $\theta_0$ (\gls{lstm} layers) and task-specific parameters $\theta_k$ (fully connected layers for task $k$). 
As opposed to layer freezing, all parameters are trainable.

@@

        \STATE \textbf{Output:} Trained multi-task \gls{lstm} metamodel $f(\cdot; \theta_0, \{\theta_k\}_{k=1}^K)$ for all $K$ tasks.
    \end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg3:multiTaskLearning} outlines the multi-task learning framework for \gls{lstm} metamodels for nested simulation procedures.
In Chapter~\ref{chap:project1} and Chapter~\ref{chap:project2}, we demonstrated that pooling using a metamodel could improve the computation efficiency and estimator accuracy in nested simulation procedures.
Here, pooling happens at a higher level, where information from multiple metamodels is shared to improve the learning performance on individual tasks.
The multi-task loss function encourages the shared \gls{lstm} layers to learn generalizable representations of temporal dependencies that are beneficial across all tasks.
